Replication on multicore considering hierarchy and characteristics of the network

### 2013-03-25 paper.tex
Haskell program to build a more or less suitable model for a very basic NUMA machine
Included first multicast try found for gruyere.

The network model is defined by communication channels, which in turn
are defined by a set of attributes (here sorted by priority):
\begin{itemize}
\item latency (quantitative) between each pair of nodes
\item ``breakdown'' (quantitative) of message send cost: Ethernet has
  dominating propagation time vs.\ dominating send and receive time on
  multicores
\item synchronized clocks: sychronized clocks + bound on propagation
  time $\rightarrow$ \emph{synchronous network} model. \\Then:
  $\exists$ algorithm for synchronous networks $\rightarrow
  \exists$ algorithm for asynchronous
\item reliability (loss, link failure, in-order delivery, phantoms,
  corruption, duplicated messages): Ethernet is not reliable,
  multicore interconnect networks often are
\item Security \& and failure detection (and the cost for that)
\end{itemize}

@@

@@

@@

\begin{itemize}
\item Need higher-level applications, such as:
  \begin{itemize}
  \item octopus
    \begin{itemize}
    \item higher level application?
    \end{itemize}
  \item capability system? \stefan{Ask Simon how expensive revocation
      currently is}
  \item a database with replication and consistency maintenance
    \begin{itemize}
    \item SharedDB? (buffers between operators, but no synchronization
      except for access to buffers, which are multiple writers, one
      reader)
    \item Crescando (some kind of state machine replication)
    \end{itemize}
  \end{itemize}
\item Machine model for topology aware tree
  \begin{itemize}
  \item Talked to Pravin about this a bit
  \end{itemize}
\item MPI collectives (as in~\cite{Tu2008})
\end{itemize}

>>>>

The network model is defined by communication channels, which in turn
are defined by a set of attributes (here sorted by priority):
\begin{itemize}
\item latency (quantitative) between each pair of nodes (per-edge)
\item ``breakdown'' (quantitative) of message send cost: Ethernet has
  dominating propagation time vs.\ dominating send and receive time on
  multicores (per-node)
\item synchronized clocks: sychronized clocks + bound on propagation
  time $\rightarrow$ \emph{synchronous network} model. \\Then:
  $\exists$ algorithm for synchronous networks $\rightarrow
  \exists$ algorithm for asynchronous
\item reliability (loss, link failure, in-order delivery, phantoms,
  corruption, duplicated messages): Ethernet is not reliable,
  multicore interconnect networks often are
\item Security \& and failure detection (and the cost for that)
\end{itemize}

@@

% Assumptions
We assume that the cost of sending a message to the same node is 0,
which is not true, since even using a highly optimized implementation
for local communication (like LMP) has a non-negligible cost.

@@

% EXAMPLE
An example of a machine model, which approximates gruyere can be found
in ..

\begin{tikzpicture}[>=latex,line join=bevel,scale=3.0]
  \pgfsetlinewidth{1bp}
\input{graphs/full_gruyere}
\end{tikzpicture}

We show the multicast tree automatically generated from a fully meshed
machine model of a 8x4x1 sample multicore (gruyere) in
Figure~\ref{fig:mst_gruyere}. We get this tree by applying a minimum
spanning tree algorithm on a multicore model encoding link latencies.

\begin{figure}
\begin{tikzpicture}[>=latex,line join=bevel,scale=.5]
  \pgfsetlinewidth{1bp}
\input{graphs/mst_gruyere}
\end{tikzpicture}
\caption{Multicast tree automatically found for a 8x4x1 multicore}
\label{fig:mst_gruyere}
\end{figure}

@@

\begin{itemize}
\item Arguing that group communication/atomic broadcast and showing
  that we can do it well might be enough (appeal to atomic broadcasts
  as the foundation for other distributed algorithms, and the MPI
  primitives, that can/need to be mapped to group-communication (have
  a list of them) 
\item Need higher-level applications, such as:
  \begin{itemize}
  \item octopus
    \begin{itemize}
    \item higher level application?
    \end{itemize}
  \item capability system? \stefan{Ask Simon how expensive revocation
      currently is}
  \item a database with replication and consistency maintenance
    \begin{itemize}
    \item SharedDB? (buffers between operators, but no synchronization
      except for access to buffers, which are multiple writers, one
      reader)
    \item Crescando (some kind of state machine replication)
    \end{itemize}
  \end{itemize}
\item Machine model for topology aware tree
  \begin{itemize}
  \item Talked to Pravin about this a bit
  \end{itemize}
\item MPI collectives (as in~\cite{Tu2008})
\end{itemize}
